### 2017-05-19 Paper Drafts/submissionDraft/submissionDraft.tex
Submitted version

\author[J.P. Meagher et al.]{J.P. Meagher, T. Damoulas, K.E. Jones, and M. Girolami\footnote{Author footnote.}}
%\index[aindx]{Author, F.} % or \aindx{Author, F.}
%\index[aindx]{Author, S.} % or \aindx{Author, S.}

\address{Department of Statistics,\\
University of Warwick, \\
J.Meagher@Warwick.ac.uk\footnote{Affiliation footnote.}}

\begin{abstract}
The reconstruction of ancestral echolocation calls is an important part of understanding the evolutionary history of bats. General techniques for the ancestral reconstruction of function-valued traits have recently been proposed. A full implementation of phylogenetic Gaussian processes for the ancestral reconstruction of function-valued traits representing bat echolocation calls is presented here. A phylogenetic signal was found in the data and ancestral reconstruction performed. This promising preliminary analysis paves the way for more realistic models of the evolution of echolocation in bats.
\end{abstract}
%\markright{Customized Running Head for Odd Page} % default is Chapter Title.
\body

@@

Previous studies of bat echolocation calls for both classification \cite{walters2012continental} and ancestral reconstruction \cite{collen2012evolution} analysed features extracted from the call spectrogram. These call features relied upon domain knowledge to ensure they were sensibly selected and applied. More recently, general techniques for the classification of acoustic signals have been developed. \cite{stathopoulos2014bat} \cite{damoulas2010bayesian} General techniques for the ancestral reconstruction of function-valued traits have also been proposed. \cite{group2012phylogenetic} Jones \& Moriarty \cite{jones2013evolutionary} extend Gaussian Process Regression \cite{rasmussen2006gaussian} to model the evolution of function-valued traits\cite{meyer2005up} over a phylogeny, a method which was demonstrated for synthetic data by Hajipantelis et al.\cite{hadjipantelis2013function}. This current research investigates these techniques in the context of the evolution of bat echolocation calls.

@@

\section{Phylogenetic Gaussian Processes}
\label{sec:pgp}
Gaussian Processes (GP) provide a concise way of fitting a prior belief over the distribution of functions. The structure of this prior belief is described by the GP model kernel, defining it's covariance function. GPs are useful for classification and regression problems, and allow posterior predictive distributions to be obtained.  An in depth treatment of GPs is given by Rasmussen \& Williams.\cite{rasmussen2006gaussian}

@@

The post processed echolocation call data accompanying Stathopoulos et al. \cite{stathopoulos2017bat} was used in this analysis. Echolocation calls were recorded across north and central Mexico with a Pettersson 1000x bat detector (Pettersson Elektronik AB, Uppsala, Sweden). Live trapped bats were measured and identified to species level using field keys.\cite{ceballos2005mamiferos} \cite{medellin2sanchez} Bats were recorded either while released from the hand or while tied to a zip line. The bat detector was set to record calls manually in real time, full spectrum at 500 kHz. 
In total the dataset consists of 22 species in five families, 449 individual bats and 1816 individual echolocation call recordings. The distribution of these call recordings across species is summarised in Table \ref{tab::dataset}.

@@

One particular limitation of the model is the representation of the echolocation call by a log energy spectral density. Echolocation calls have complex spectral and temporal structures, much of which is lost in the log energy spectral density representation. An alternative representation, which preserves more of this structure, is the spectrogram. Implementing this model using the spectrogram as a functional representation of the echolocation call is a priority, as the results of such an analysis are expected to be more meaningful.

>>>>

\author[J.P. Meagher et al.]{J.P. Meagher\footnote{Department of Statistics, The University of Warwick}, T. Damoulas\footnote{The University of Warwick and The Alan Turing Institute}, K.E. Jones\footnote{Centre for Biodiversity and Environment Research, University College London}, and M. Girolami\footnote{Imperial College London and The Alan Turing Institute}\\}
%\index[aindx]{Author, F.} % or \aindx{Author, F.}
%\index[aindx]{Author, S.} % or \aindx{Author, S.}

\address{
J.Meagher@Warwick.ac.uk}

\begin{abstract}
The reconstruction of ancestral echolocation calls is an important part of understanding the evolutionary history of bats. General techniques for the ancestral reconstruction of function-valued traits have recently been proposed. A full implementation of phylogenetic Gaussian processes for the ancestral reconstruction of function-valued traits representing bat echolocation calls is presented here. A phylogenetic signal was found in the data and ancestral reconstruction performed. This promising preliminary analysis paves the way for more realistic models for the evolution of echolocation in bats.
\end{abstract}
%\markright{Customized Running Head for Odd Page} % default is Chapter Title.
\body

@@

Previous studies of bat echolocation calls for both classification \cite{walters2012continental} and ancestral reconstruction \cite{collen2012evolution} analysed features extracted from the call spectrogram. These call features relied upon domain knowledge to ensure they were sensibly selected and applied. More recently, general techniques for the classification of acoustic signals have been developed. \cite{stathopoulos2014bat} \cite{damoulas2010bayesian} General techniques for the ancestral reconstruction of function-valued traits have also been proposed. \cite{group2012phylogenetic} Jones \& Moriarty \cite{jones2013evolutionary} extend Gaussian Process Regression \cite{rasmussen2006gaussian} to model the evolution of function-valued traits\cite{meyer2005up} over a phylogeny, a method which was demonstrated for synthetic data by Hajipantelis et al.\cite{hadjipantelis2013function}. This current research investigates these techniques in the context of the bat echolocation calls.

@@

\section{Phylogenetic Gaussian Processes}
\label{sec:pgp}
A Gaussian process places a prior distribution over functions, \(f(x) \sim \mathcal{GP}(m(x), k (x, x'),\) where \(x \in \mathbf{R}^P\) is some input variable, the mean function \(m(x) = \mathbf{E}[f(x)]\), and the covariance function \(k(x, x') = \mathnormal{cov}(f(x), f(x') )\). Given observations \(\mathbf{y}\) at locations \(\{x_n\}_{n=1}^N\), Gaussian noise, and kernel hyperparameters \(\theta\), a posterior predictive distribution over functions can be inferred analytically. See Rasmussen \& Williams\cite{rasmussen2006gaussian} for an in depth treatment.

@@

The post processed echolocation call data for bats found across north and central Mexico which accompanied Stathopoulos et al. \cite{stathopoulos2017bat} was used in this analysis. Live bats were caught, identified, and recorded either while released from the hand or while tied to a zip line. Calls were recorded at a sampling frequency of 500 kHz. 
In total the dataset consists of 22 species in five families, 449 individual bats and 1816 individual echolocation call recordings. The distribution of these call recordings across species is summarised in Table \ref{tab::dataset}.

@@

One particular limitation of the model is the representation of the echolocation call by a log energy spectral density. Echolocation calls have complex spectral and temporal structures, much of which is lost in the log energy spectral density representation. An alternative time-frequency representation, which preserves more of this structure, is the spectrogram. Modelling the evolution of bat echolocation calls with spectrograms is an active area of future research.
